# QFASA

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```


# Overview
Accurate estimates of the diets of predators are required in many
areas of ecology, but for many species current methods are imprecise,
limited to the last meal, and often biased. The diversity of fatty
acids and their patterns in organisms, coupled with the narrow
limitations on their biosynthesis, properties of digestion in
monogastric animals, and the prevalence of large storage reservoirs of
lipid in many predators, led us to propose the use of quantitative
fatty acid signature analysis (QFASA) to study predator diets.

This package allows diet estimates to be calculated via QFASA, MUFASA 
(or MLE QFASA) and by simultaneous estimation of calibration coefficients 
and diets.  It also contains functions to compute confidence intervals,
measures of repeatability, pseudo predator generation, etc...


# Installing
## Via GitHub

```r
devtools::install_github('cstewartGH/QFASA')
```

## Via CRAN

```r
install.packages('QFASA')
```

# Load Package

```{r, eval=TRUE}
library(QFASA)
```

# Vignettes

Please refer to vignettes for detailed instructions on how to set up data
in order to estimate the diets of predators via QFASA or MUFASA.


