---
title: "Modeling Workflow"
author: "Shelly Lang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


# Modeling Inputs
Prior to starting make sure that:

* Fatty acid names in all files are the same (contain the exact same
  numbers/characters and punctuation)
* There are no fatty acids in the prey file that do not appear in the
  predator file and visa versa


## Distance Measure
Choose from one of three distance measures:  
1=KL (Kullback-Leibler) || 2=AIT (Aitchison) || 3=CSD (Chi-Squared) 

```{r}
dist.meas=1
```

## Fatty Acid Set
* This is the list of FAs to be used in the modelling.
* The simplest alternative is to load a .csv file which contains a
  single column with a header row and the names of the fatty acids
  listed below (see example file __"FAset.csv"__).
* A more complicated alternative is to load a .csv file with the full
  set of FAs and then add code to subset the FAs you wish to use from
  that set --> this alternative is useful if you are planning to test
  multiple FA sets.
* Regardless of how you load the FA set it must be converted to a
  vector.

```{r, eval=FALSE}
fa.set=read.csv(file="FAset.csv",as.is=TRUE)
fa.set=as.vector(unlist(fa.set))
```

## Matrix of Predator FA signatures
* the FA signatures in the originating .csv file should be in percentages not proportions (i.e. adding to ~100 not 1)  
* each predator signature is a row with the FAs in columns (see
  example file __"predatorFAs.csv"__)
* the FA signatures are subsetted for the chosen FA set (created above) and renormalized during the modelling so there is no need to subset and/or renormalize prior to loading the .csv file or running p.QFASA  
    + BUT make sure that the the same FAs appear in the predator and
      prey files (if a FA appears in one but not the other the code
      will give you an error)
* unlike the original QFASApack code the predator FA .csv file can
  contain as much tombstone data in columns as you wish but the
  predator FA signatures must be extracted as a separate input in
  order to run in p.QFASA
    + For example: in the code below the predator .csv file
      ("predatorFAs.csv") has 4 tombstone columns (SampleCode,
      AnimalCode, SampleGroup, Biopsy). Prior to running QFASA the
      tombstone (columns 1-4) and FA data (columns 5 onward) are each
      extracted from the original data frame. The FA data becaomes the
      the predator.matrix (which is passed to p.QFASA) and the
      tombstone data is retained so that it can be recombined with the
      model output later on.


## Matrix of Prey FA signatures
* The FA signatures in the originating .csv file should be in
  percentages not proportions (i.e. adding to ~100 not 1).
* The prey file should contain all of the individual FA signatures of
  the prey and their lipid contents (where appropriate) - a matrix of
  the mean values for the FAs (prey.matrix) by the designated prey
  modelling group is then calculated using the MEANmeth function
  loaded above.
* Like the predator .csv file you can have as many tombstone data
  columns as required but there must be at least one column that
  identifies the modelling group to be used (in the example file used
  below __"preyFAs.csv"__ it is the "Species" column)
* unlike the predator data, the prey data is not subsetted and
  renomalized during the modelling so the prey file needs to be
  subsetted for the desired FA set (created above) and renormalized to
  sum to 1 prior to calculating the mean values (see code below)
    + example: in the code below the "preyFAs.csv" file has 3
      tombstone columns. The full FA set is extracted from the data
      frame (columns 4 onward), subsetted for the FA set in use and
      then renormalized over 1. The modelling group names (the
      "Species" column in this case) is then added back to the
      subsetted and renormalized data (as the first column) and the
      average values calculated using the MEANmeth function. Note that
      for the MEANmeth function to work the modelling group name must
      be in the first column
    
```{r, eval=FALSE }
#full file
prey=read.csv(file="preyFAs.csv", as.is=TRUE)
#extract prey FA only from data frame and subset them for the FA set designated above
prey.sub=(prey[,4:(ncol(prey))])[fa.set]
#renormalize over 1
prey.sub=prey.sub/apply(prey.sub,1,sum) 
#extract the modelling group names from the full file
group=as.vector(prey$Species) 
#add modelling group names to the subsetted and renormalized FAs
prey.matrix=cbind(group,prey.sub)
#create an average value for the FA signature for each designated modelling group
prey.matrix=MEANmeth(prey.matrix) 
```

## Prey lipid content
* mean lipid content by modelling group is calculated from the full
  prey file using the modelling group as a summary variable (see code
  below).
* **Note:** if no lipid content correction is going to be applied then
  a vector of '1's of length equal to the number of modelling groups
  is used as the vector instead i.e. FC=rep(1,nrow(prey.matrix))

```{r, eval=FALSE }
FC=prey[,c(2,3)] #numbers are the column which identifies the modelling group, and the column which contains the lipid contents
FC=as.vector(tapply(FC$lipid,FC$Species,mean,na.rm=TRUE))
```


## Calibration Coefficients
* Originating .csv file should contain 2 columns (with headers). The
  first contains the FA names, the second the value of the CC for each
  FA (see example file __"CC.csv"__).
* __IMPORTANT:__ the FAs in the CC.csv file __MUST__ be exactly the
  same as the FAs in the originating predator.csv file __AND__ they
  __MUST__ BE IN THE __*EXACT*__ SAME ORDER.
  
```{r, eval=FALSE}
cal=read.csv(file="CC.csv", as.is=TRUE)
cal.vec=cal[,2]
cal.mat=replicate(npredators,cal.vec)
```


# Run QFASA

